---
title: "Arima_Assignment"
output: html_document
date: "2023-11-22"
---

### To Load the Data and update the column names.
```{r}
library(readxl)
MyDf=read_excel("TimeSeriesData_LiquourSales.xls")

colnames(MyDf) <- c("Date", "Sales")

MyDf$Date <- as.Date(MyDf$Date)
```


### Converting the Data into Time series and plotting it along with ACF.
```{r}
library(fpp)
library(fpp2)
library(forecast)

#Converting data into TimeSeries:

Tsales <- ts(MyDf$Sales, start = c(2018,1), end = c(2023,7), frequency = 12)

plot(Tsales)

Acf(Tsales)
```

## To check if Time series is stationary.

#### ADF test says differences is required if p-value is > 0.05. This means we need to try differencing the time series to achieve stationarity.

```{r}
adf.test(Tsales)
```


### Since this Time series is seasonal, we will use function nsdiffs.
#### NSDIFFS = 1 means differencing the series by subtracting each observation from the corresponding observation in the previous year. 
```{r}
nsdiffs(Tsales)
ndiffs(Tsales) #For non-seasonal data()

# tsdisplay plots ACF,PACF and timeseries plot together.

tsdisplay(Tsales)

timeseriesdiff1 <- diff(Tsales, differences=1)
plot(timeseriesdiff1)

```

### To check if more difference is needed:

```{r}
tsdisplay(timeseriesdiff1)
```

### Using Auto-Arima Model and plotting it.


```{r}
auto_fit <- auto.arima(Tsales, trace=TRUE, stepwise = FALSE)

auto_fit
attributes(auto_fit)
```
#### Best model: ARIMA(0,1,2)(0,1,1)[12] 

##### Explanation: The notation "ARIMA(0,1,2)(0,1,1)[12]" represents a specific type of time series forecasting model in the context of ARIMA (AutoRegressive Integrated Moving Average) models with seasonality. Let's break down each component:

##### ARIMA(p, d, q): 
###### p: Order of the autoregressive (AR) component.
###### d: Degree of differencing (non-seasonal).
###### q: Order of the moving average (MA) component.
###### In the notation "ARIMA(0,1,2)", it means there is no autoregressive component (p=0), first-order differencing is applied (d=1), and there are two lagged moving average terms (q=2). This implies that the model includes two lagged values of the differenced series in the moving average part.

##### Seasonal Component (SARIMA):

###### (P, D, Q)[s]: Seasonal order components.
###### P: Seasonal autoregressive order.
###### D: Seasonal degree of differencing.
###### Q: Seasonal moving average order.
###### s: Seasonal period (e.g., 12 for monthly data).
###### In the notation "(0,1,1)[12]", it means there is no seasonal autoregressive component (P=0), first-order seasonal differencing is applied (D=1), one lagged value of the seasonally differenced series is included in the seasonal moving average part (Q=1), and the seasonal period is 12 (indicating monthly data).

##### Putting it all together, "ARIMA(0,1,2)(0,1,1)[12]" represents a seasonal ARIMA model where first-order differencing is applied both non-seasonally and seasonally, and the model includes lagged moving average terms in both the non-seasonal and seasonal components. 


### Plot the forecast for next 5 years

```{r}
F_Arima<-forecast(auto_fit,h=5,level=c(99.5))

plot(F_Arima)
```

### Residual Analysis
```{r}
Acf(auto_fit$residuals)
plot.ts(residuals(auto_fit))
hist(auto_fit$residuals)
tsdiag(auto_fit)
```
##### ACF- All the points are below significant line, so they dont have any correaltion.
##### Plot- There is no significant pattern shown between the residual.
##### Histogram- It is also, normaly distributed, mean is close to zero which is a good thing.
 
### Accuracy For ARIMA model

##### MAPE is approximately 2.0483%, providing a percentage measure of accuracy.
```{r}
accuracy(F_Arima)
```

